%!TEX root = thesis.tex
This thesis would have not been possible without the support from my advisor, Prof. Enrico Bertini. Thank you for the numerous discussions about my research, being there whenever I had doubts, supporting my good ideas, and dismissing my bad ideas.
You are an awesome advisor and I would do my Ph.D. again under your supervision.
I thank our collaborators and co-authors for the exciting paths we ventured.
Thank you Adam Perer, Aritra Dasgupta, Yindalon Aphinyanaphongs, Jordan Swartz, Kenney Ng, and many more for your advice and inputs throughout our collaborations.
I would also like to extend my gratitude to Prof. Cl\'audio Silva, Prof. Juliana Freire, Prof. Rumi Chunara, and Prof. Lisa Hellerstein for their guidance, experience, and feedback at various stages of my Ph.D.

Thanks, Adam and Aritra, for the amazing internships at both IBM and PNNL. I learned a lot.
A special thanks goes to my lab-mates, Anshul, Cristian, Jay, and Jun for their constant support, feedback, and always having an open ear for discussions.
Also, I thank our interns Paolo, Livia, and Yao. I wish you all the best.
Furthermore, I thank every current or former member of the VIDA family with whom I spent many hours and made so many friends. Thank you R\'emi, Vicky, Fernando, Fabio, Kien, Tuan-Anh, Harish, Aline, Bowen, Sungmin, Gromit, Masayo, Yamuna, Raoni, Jorge, Felipe, Aecio, Sonia, and so many others whom I might have missed to mention.
I extend my warmest thanks to all the administrators who stayed on top of the administrative processes, making sure that I can focus on my research without any distractions. Thank you Ann, Susana, Judy, Kari, and Eve for your support all these years.

Additionally, I thank Hendrik for encouraging and supporting the best decision of my life, and Narges for being herself.
Also, I thank Cl\'ement, Nandini, Jeremy, Pauline, Daniel, Leonard, Ingrid and her family, my Volleyball team, Jason, Keishi, Michael, Pavol, Veronica, Dunia, Cerille, Thierry, and many others for fun times and keeping my sanity during the years of my Ph.D.
On a personal note, I thank my parents for their love and support throughout my life.
And my siblings, Vanessa, Michaela, Swen, and adjunct siblings, for showing me the important things in life.
You are the best.

\noindent
\makebox[\textwidth]{\hfill\makebox[3in]{\hfill Josua Krause\hfill}}
\makebox[\textwidth]{\hfill\makebox[3in]{\hfill\graddate\hfill}}
